---
title: "World Happiness EDA"
author: "Avery Rogers"
date: '2021-03-26'
output:
  html_document:
    df_print: paged
slug: world-happiness-eda
---

**TL;DR** According to data from the 2021 World Happiness Report, wealth, social support, and life expectancy all seem to correlate strongly with happiness across countries. Perceptions of corruption appears to correlate negatively with happiness, though outliers exist. Generosity does not seem to correlate with happiness at all, though the proxy for generosity (donating to a charity within the last month) seems ill-adapted to capture the kinds of informal help offered between members of distinct communities that dominate poorer and politically fragile countries. 

*Code can be found on Github [right here.](https://github.com/averyr2/r_averyblog/blob/master/content/post/2021-03-26-world-happiness-EDA/index.Rmd)

------------------------------------------------------------------------------

Today, I will use Exploratory Data Analysis (EDA) using R's `dplyr` and `ggplot2` package suites to investigate the correlates of happiness across countries. I used the World Happiness Report for all measures related to happiness, and the World Bank 2019 population data for country populations.

*Links to datasets*:
[World Happiness Report 2021 Data](https://www.kaggle.com/mathurinache/world-happiness-report-2021),
[World Bank 2019 Data](https://data.worldbank.org/indicator/SP.POP.TOTL)

After cleaning and merging the data, I plotted per capita GDP, life expectancy, social support, generosity, and corruption as they relate to happiness. I sized each circle in proportion to the country's population, with India and China being the two largest by a wide margin. (Note that all smooth curves are calculated using local polynomial regression, and that they are weighted equally by country and not by population.)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = FALSE, warning = FALSE)
```

```{r results=FALSE, warning=FALSE, message=FALSE}
# Libraries
library(tidyverse)
library(ggrepel)

# Parameters
current_ladders_data <- "~/Downloads/world_happiness/current_ladders.csv"
historical_ladders_data <- "~/Downloads/world_happiness/historical_ladders.csv"
current_mortality_data <- "~/Downloads/world_happiness/current_mortality.csv"
world_pop_data <- "~/Downloads/world_happiness/world_pop.csv"
#===============================================================================

# Generic functions for data parsing
divide_1000 <- function(x, na.rm=FALSE) {
  x / 1000 
}

make_decimal <- function(x, na.rm=FALSE) {
  as.integer(str_sub(x, 3, -1)) / 1000 
}

```

```{r results=FALSE, warning=FALSE, message=FALSE}
current_ladders <- read_csv(current_ladders_data)

current_ladders <- 
  current_ladders %>% 
  rename(
    country = `Country name`,
    region = `Regional indicator`,
    ladder_score = `Ladder score`,
    std_error_ladder_score = `Standard error of ladder score`,
    log_gdp_pc = `Logged GDP per capita`,
    social_support = `Social support`,
    healthy_life_expectancy = `Healthy life expectancy`,
    freedom_of_choice = `Freedom to make life choices`,
    generosity = Generosity,
    perceptions_corruption = `Perceptions of corruption`,
    ladder_dystopia = `Ladder score in Dystopia`,
    explained_log_gdp_pc = `Explained by: Log GDP per capita`,
    explained_social_support = `Explained by: Social support`,
    explained_life_expectancy = `Explained by: Healthy life expectancy`,
    explained_freedom_choices = `Explained by: Freedom to make life choices`,
    explained_generosity = `Explained by: Generosity`,
    explained_corruption = `Explained by: Perceptions of corruption`,
    dystopia_residual = `Dystopia + residual`
  ) %>% 
  mutate_at(
    vars(social_support, freedom_of_choice, generosity, perceptions_corruption), 
    make_decimal
    ) %>% 
  mutate_at(
    vars(
      ladder_score, 
      log_gdp_pc,
      healthy_life_expectancy
    ), 
    divide_1000
  )

current_ladders

```

```{r results=FALSE, warning=FALSE, message=FALSE}
world_pop <- 
  read_csv(world_pop_data, skip = 4) %>% 
  select(country = `Country Name`, pop_2019 = `2019`) 

world_pop$country <- 
  world_pop$country %>% 
  # making sure names match those in Happiness database
  recode(
    `Congo, Rep.` = "Congo (Brazzaville)", 
    `Egypt, Arab Rep.` = "Egypt",
    `Gambia, The` = "Gambia",
    `Hong Kong SAR, China` = "Hong Kong S.A.R. of China",
    `Iran, Islamic Rep.` = "Iran",
    `Cote d'Ivoire` = "Ivory Coast",
    `Kyrgyz Republic` = "Kyrgyzstan",
    `Lao PDR` = "Laos",
    `Russian Federation` = "Russia",
    `Slovak Republic` = "Slovakia",
    `Korea, Rep.` = "South Korea",
    `Eswatini` = "Swaziland",
    `Venezuela, RB` = "Venezuela",
    `Yemen, Rep.` = "Yemen"
  )

world_pop 
```

```{r}
current_ladders <- 
  current_ladders %>% 
  left_join(world_pop, by = "country") 

current_ladders %>% 
  arrange(country)
```

```{r}
current_ladders$region <- 
  current_ladders$region %>% 
  recode(
    `Western Europe` = "Europe", 
    `Central and Eastern Europe` = "Europe",
    `East Asia` = "Asia",
    `Southeast Asia` = "Asia",
    `Commonwealth of Independent States` = "Commonwealth of\nIndependent States",
    `North America and ANZ` = "North America, Australia,\nand New Zealand"
  )
```


```{r results=FALSE, warning=FALSE, message=FALSE, width = 6}
current_ladders %>% 
  arrange(desc(pop_2019)) %>% 
  ggplot(aes(log_gdp_pc, ladder_score)) + 
  geom_smooth() +
  geom_point(aes(fill = region, size = pop_2019), pch = 21) +
  scale_fill_brewer(palette = "RdYlBu") +
  scale_radius(range = c(2, 16)) +
  scale_y_continuous(
    limits = c(2, 8.1), 
    breaks = c(3, 4, 5, 6, 7, 8), 
    expand = c(0, 0)
  ) +
  annotate("text", label = "United States", x = 10, y = 7.75) +
  geom_segment(aes(x = 10, y = 7.55, xend = 11.05, yend = 6.95), size= 0.1) +
  annotate("text", label = "China", x = 10.5, y = 4.45) +
  geom_segment(aes(x = 10.5, y = 4.6, xend = 9.8, yend = 5.35), size= 0.1) +
  annotate("text", label = "India", x = 8.9, y = 3.2) +
  labs(
    x = "Log GDP Per Capita",
    y = "Happiness Ladder Score",
    title = "Wealthier Countries are Happier on Average",
    fill = "Region",
    subtitle = "Higher GDP per capita correlates with higher happiness scores"
  ) +
  guides(size = FALSE)
``` 

As we can see in the plot above, happiness seems to increase pretty steadily with per capita GDP. We also notice, unsurprisingly, strong regional patterns: Europe and North America come out with both high per capita GDP and happiness scores, while many sub-Saharan African countries come out with low per capita GDP and happiness scores. India is interestingly below our smooth curve, indicating that it is less happy than its per capita GDP peers, at least according to the dataset I've used. 

```{r, results=FALSE, warning=FALSE, message=FALSE, width = 6}
current_ladders %>% 
  filter(region == "Latin America and Caribbean") %>% 
  filter(ladder_score == max(ladder_score))
```

On a cute sidenote, that yellow dot that's well above the rest of Latin America and the Caribbean is Costa Rica. I traveled there as a teenager and it was indeed one of the most beautiful, laid back, and positive places I've ever been. I'm glad that wasn't just a naïve impression, at least according to the data on hand! 


```{r results=FALSE, warning=FALSE, message=FALSE, width = 6}
current_ladders %>% 
  arrange(desc(pop_2019)) %>% 
  ggplot(aes(healthy_life_expectancy, ladder_score)) + 
  geom_smooth() +
  geom_point(aes(fill = region, size = pop_2019), pch = 21) +
  scale_fill_brewer(palette = "RdYlBu") +
  scale_radius(range = c(2, 16)) +
  scale_y_continuous(
    limits = c(2, 8.1), 
    breaks = c(3, 4, 5, 6, 7, 8), 
    expand = c(0, 0)
  ) +
  scale_x_continuous(
    limits = c(47, 78),
    breaks = c(50, 55, 60, 65, 70, 75), 
    expand = c(0, 0)
  ) +
  annotate("text", label = "United States", x = 65, y = 7.3) +
  annotate("text", label = "China", x = 71, y = 4.75) +
  annotate("text", label = "India", x = 63.7, y = 3.7) +
  labs(
    x = "Healthy Life Expectancy",
    y = "Happiness Ladder Score",
    title = "Healthier Countries are Happier on Average",
    fill = "Region",
    subtitle = "Higher healthy life expectancy correlates with higher happiness scores"
  ) +
  guides(size = FALSE)
```

We see a similar curve for healthy life expectancy as we do for per capita GDP: countries with longer healthy life expectancies are happier on average. (Note that healthy life expectancy means life expectancy adjusted for the suffering brought upon by significant disease and disability). Interestingly here, we see large parts of Europe falling well above the curve compared to their Asian counterparts with similar life expectancies, indicating that other geographically-relevant cultural patterns are at play in differentiating the happiness of those countries with high health.  

```{r results=FALSE, warning=FALSE, message=FALSE, width = 6}
current_ladders %>% 
  arrange(desc(pop_2019)) %>%
  ggplot(aes(social_support, ladder_score)) + 
  geom_smooth() +
  geom_point(aes(fill = region, size = pop_2019), pch = 21) +
  scale_fill_brewer(palette = "RdYlBu") +
  scale_radius(range = c(2, 16)) +
  scale_y_continuous(
    limits = c(2, 8.1), 
    breaks = c(3, 4, 5, 6, 7, 8), 
    expand = c(0, 0)
  ) +
  scale_x_continuous(limits = c(0.45, 1.0), expand = c(0, 0)) +
  geom_segment(
    aes(x = 0.805, y = 2.4, xend = 0.915, yend = 2.4), 
    arrow = arrow(length = unit(0.15, "cm"), type = "closed") 
  ) + 
  annotate(
    "text", 
    label = "Higher social\n support", 
    x = 0.85, 
    y = 2.8, 
    size = 3, 
    colour = "black"
  ) +
  geom_segment(
    aes(x = 0.645, y = 2.4, xend = 0.535, yend = 2.4), 
    arrow = arrow(length = unit(0.15, "cm"), type = "closed") 
  ) + 
  annotate(
    "text", 
    label = "Lower social\n support", 
    x = 0.585, 
    y = 2.8, 
    size = 3, 
    colour = "black"
  ) +
  annotate("text", label = "United States", x = 0.84, y = 7.6) +
  geom_segment(aes(x = 0.84, y = 7.45, xend = 0.92, yend = 6.95), size= 0.1) +
  annotate("text", label = "China", x = 0.7, y = 6) +
  geom_segment(aes(x = 0.7, y = 5.85, xend = 0.8, yend = 5.5), size= 0.1) +
  annotate("text", label = "India", x = 0.655, y = 3.6) +
  labs(
    x = "Social Support (Self-Reported)",
    y = "Happiness Ladder Score",
    title = "More Socially Supportive Countries are Happier on Average",
    fill = "Region",
    subtitle = "Self-reported levels of social support correlate with higher happiness scores"
  ) +
  guides(size = FALSE)
```

In the World Happiness Survey, social support is quantified by a binary variable corresponding to the following question: "If you were in trouble, do you have relatives or friends you can count on to help you whenever you need them, or not?". The average score is obtained for each country between 0 (no social support) and 1 (high social support). We see here that many countries in Europe, North America, Australia, and New Zealand, Latin America and the Caribbean, and Asia report very high levels of social support. India is an interesting outlier in our diagram with very low reported levels of social support; it would be an interesting extension of this work to try and understand the factors influencing social connectedness in India.

```{r results=FALSE, warning=FALSE, message=FALSE}
current_ladders %>% 
  filter(social_support == min(social_support))

current_ladders %>% 
  filter(region == "South Asia") %>% 
  select(country, social_support)

```


The country with the least social support on our graph is Afghanistan, with an average social support score of only .463. Neighboring Pakistan has a social support level of .651, which is higher than India at .603. 


```{r results=FALSE, warning=FALSE, message=FALSE, width = 6}
current_ladders %>% 
  filter(country != "Afghanistan") %>% 
  arrange(desc(pop_2019)) %>%
  ggplot(aes(freedom_of_choice, ladder_score)) + 
  geom_smooth() + 
  geom_point(aes(fill = region, size = pop_2019), pch = 21) +
  scale_fill_brewer(palette = "RdYlBu") +
  scale_radius(range = c(2, 16)) +
  scale_y_continuous(breaks = c(3, 4, 5, 6, 7, 8)) +
  geom_segment(
    aes(x = 0.78, y = 2.4, xend = 0.89, yend = 2.4), 
    arrow = arrow(length = unit(0.15, "cm"), type = "closed") 
  ) + 
  annotate(
    "text", 
    label = "Higher Freedom\nof Choice", 
    x = 0.84, 
    y = 2.8, 
    size = 3, 
    colour = "black"
  ) +
  geom_segment(
    aes(x = 0.65, y = 2.4, xend = 0.54, yend = 2.4), 
    arrow = arrow(length = unit(0.15, "cm"), type = "closed") 
  ) + 
  annotate(
    "text", 
    label = "Lower Freedom\nof Choice", 
    x = 0.6, 
    y = 2.8, 
    size = 3, 
    colour = "black"
  ) +
  annotate("text", label = "United States", x = 0.8, y = 7.6) +
  geom_segment(aes(x = 0.8, y = 7.45, xend = 0.84, yend = 6.95), size= 0.1) +
  annotate("text", label = "China", x = 0.94, y = 4.6) +
  geom_segment(aes(x = 0.94, y = 4.75, xend = 0.91, yend = 5.3), size= 0.1) +
  annotate("text", label = "India", x = 0.945, y = 3.6) +
  labs(
    x = "Freedom of Choice (Self-Reported)",
    y = "Happiness Ladder Score",
    title = "Countries with more Personal Freedoms are Happier on Average",
    fill = "Region",
    subtitle = "Higher self-reported freedom of choice correlates with higher happiness scores"
  ) +
  guides(size = FALSE)
```

Like social support, the freedom of choice metric is a national average over the binary question: "Are you satisfied or dissatisfied with your freedom to choose what you do with your life?". 

For this part, I actually had to remove Afghanistan because its freedom of choice score was so low - less than 0.4 - that it threw of the local regression curve. It's not a great time to be Afghani, unfortunately.

Otherwise, we see a gentle but noisy increase in happiness as freedom of choice expands for people. There does not seem to be as much geographic clustering for freedom of choice as there was for the other metrics we've looked at. Within Europe, for example, Norway has the highest freedom of choice score of 0.96, while Greece comes in last with a score of 0.562. That seems somewhat unsurprising given the poor financial and employment outcomes in Greece compared to its Northern and Western European counterparts. 


```{r}
current_ladders %>% 
  filter(freedom_of_choice == min(freedom_of_choice))

current_ladders %>%
  filter(region == "Europe") %>% 
  select(country, freedom_of_choice) %>% 
  arrange(desc(freedom_of_choice))
```


```{r results=FALSE, warning=FALSE, message=FALSE, width = 6}
current_ladders %>% 
  arrange(desc(pop_2019)) %>%
  ggplot(aes(generosity, ladder_score)) + 
  geom_smooth() +
  geom_point(aes(fill = region, size = pop_2019), pch = 21) +
  scale_fill_brewer(palette = "RdYlBu") +
  scale_radius(range = c(2, 16)) +
  scale_y_continuous(breaks = c(3, 4, 5, 6, 7, 8)) +
  geom_segment(
    aes(x = 0.34, y = 2.4, xend = 0.45, yend = 2.4), 
    arrow = arrow(length = unit(0.15, "cm"), type = "closed") 
  ) + 
  annotate(
    "text", 
    label = "More Generous", 
    x = 0.39, 
    y = 2.7, 
    size = 3, 
    colour = "black"
  ) +
  geom_segment(
    aes(x = 0.18, y = 2.4, xend = 0.07, yend = 2.4), 
    arrow = arrow(length = unit(0.15, "cm"), type = "closed") 
  ) + 
  annotate(
    "text", 
    label = "Less Generous", 
    x = 0.13, 
    y = 2.7, 
    size = 3, 
    colour = "black"
  ) +
  annotate("text", label = "United States", x = 0.2, y = 8) +
  geom_segment(aes(x = 0.13, y = 7.8, xend = 0.1, yend = 6.95), size= 0.1) +
  annotate("text", label = "India", x = 0.145, y = 4.1) +
  labs(
    x = "Generosity (Self-Reported)",
    y = "Happiness Ladder Score",
    title = "More Generous Countries are No Happier on Average",
    fill = "Region",
    subtitle = "Nationwide generosity levels do not correlate with higher happiness scores"
  ) +
  guides(size = FALSE)
```

We now turn to a variable that does not seem to have much to do with happiness across countries: generosity. In the World Happiness Report, generosity is a national average across the binary question: "Have you donated money to a charity in the past month?". 

We see fewer country responses with the majority of countries having a score below 0.2. As such, this seems like a poor metric for evaluating differential generosity if all countries are clustered in a narrow band near 0. It also seems like a poor metric because it does not get at the heart of interpersonal generosity that dominates many communities in an informal manner. 

```{r}
current_ladders %>% 
  top_n(5, generosity) %>% 
  select(country, generosity)
```


The top 5 countries for generosity are fascinating: 

1. Indonesia, with a score of 0.542. 
2. Myanmar, with a score of 0.509. 
3. Gambia, with a score of 0.424. 
4. Haiti, with a score of 0.422. 
5. Uzbekistan, with a score of 0.311.

I have no particular explanation for this, but I'd be curious to find one!


```{r results=FALSE, warning=FALSE, message=FALSE, width = 6}
current_ladders %>% 
  filter(country != "Rwanda") %>% 
  arrange(desc(pop_2019)) %>%
  ggplot(aes(perceptions_corruption, ladder_score)) + 
  geom_smooth() +
  geom_point(aes(fill = region, size = pop_2019), pch = 21) +
  scale_fill_brewer(palette = "RdYlBu") +
  scale_radius(range = c(2, 16)) +
  scale_y_continuous(
    breaks = c(3, 4, 5, 6, 7, 8), 
    limits = c(2.3, 8), 
    expand = c(0,0)
  ) +
  geom_segment(
    aes(x = 0.6, y = 2.5, xend = 0.76, yend = 2.5), 
    arrow = arrow(length = unit(0.15, "cm"), type = "closed") 
  ) + 
  annotate(
    "text", 
    label = "More Corrupt", 
    x = 0.68, 
    y = 2.8, 
    size = 3, 
    colour = "black"
  ) +
  geom_segment(
    aes(x = 0.4, y = 2.5, xend = 0.24, yend = 2.5), 
    arrow = arrow(length = unit(0.15, "cm"), type = "closed") 
  ) + 
  annotate(
    "text", 
    label = "Less Corrupt", 
    x = 0.33, 
    y = 2.8, 
    size = 3, 
    colour = "black"
  ) +
  annotate("text", label = "United States", x = 0.8, y = 7.6) +
  geom_segment(aes(x = 0.77, y = 7.45, xend = 0.7, yend = 6.95), size= 0.1) +
  annotate("text", label = "China", x = 0.57, y = 4.6) +
  geom_segment(aes(x = 0.57, y = 4.75, xend = 0.76, yend = 5.5), size= 0.1) +
  annotate("text", label = "India", x = 0.93, y = 4.1) +
  geom_segment(aes(x = 0.93, y = 3.95, xend = 0.77, yend = 3.8), size= 0.1) +
  labs(
    x = "Perceptions of Corruption (Self-Reported)",
    y = "Happiness Ladder Score",
    title = "More Corrupt Countries are Less Happy on Average",
    fill = "Region",
    subtitle = "Perceived corruption correlates with lower happiness scores"
  ) +
  guides(size = FALSE)
```

Finally, we have perceptions of corruption. Perceptions of corruption are measured using two questions in the World Happiness Survey: "Is corruption widespread throughout the government or not?" and "Is corruption widespread within businesses or not?". The answers are averaged by person and a national average is taken over the answers. (Note that these scores do not necessarily reflect actual corruption, which, by definition, is unmeasurable unless it is caught). 

```{r}
current_ladders %>% 
  filter(perceptions_corruption < 0.25) %>% 
  filter(ladder_score < 5) %>% 
  select(country, ladder_score, perceptions_corruption)
```

Here, perceived corruption is greater on the righthand side of the graph. We have one extreme outlier, Rwanda, which has a low ladder score of 3.415 but also a very low perception of corruption, 0.167. I removed Rwanda for the sake of seeing how our local regression curve looks without this outlier. Once Rwanda has been removed, we see that countries with lower corruption scores (mostly in Europe and North America/ANZ) tend to be happier. We also notice that our geographic segmentation is back, with sub-Saharan African countries reporting high corruption (other than Rwanda!) and low happiness while European countries with high corruption are generally pretty happy. 

```{r}
current_ladders %>% 
  filter(perceptions_corruption == min(perceptions_corruption)) %>% 
  select(country, perceptions_corruption)
```


The country with the least perceived corruption is Singapore, with a perceptions of corruption score of only 0.082. I suppose all those strict rules create a strong social fabric of trust! 

----------------------------------

All in all, we see that wealth, healthy long lives, and social support are strong correlates of happiness (not too surprising!) and that corruption is inversely correlated with happiness. While this particular dataset offers no conclusions about generosity and happiness, I imagine a more robust measure of generosity beyond the current binary question would reveal interesting results. 


```{r results=FALSE, warning=FALSE, message=FALSE}
# Everything from here down is code that I did not use in my blog post but am 
# currently toying with. It may go live, so I'll leave it here for safe keeping!

historical_ladders <- read_csv(historical_ladders_data)

historical_ladders <- 
  historical_ladders %>% 
  select(
    country = `Country name`,
    year,
    ladder_score = `Life Ladder`,
    log_gdp_pc = `Log GDP per capita`,
    social_support = `Social support`,
    healthy_life_expectancy = `Healthy life expectancy at birth`,
    freedom_of_choice = `Freedom to make life choices`,
    generosity = Generosity,
    perceptions_corruption = `Perceptions of corruption`
  ) %>% 
  mutate_at(
    vars(social_support, freedom_of_choice, generosity, perceptions_corruption), 
    make_decimal
    ) %>% 
  mutate_at(
    vars(
      ladder_score, 
      log_gdp_pc,
      healthy_life_expectancy
    ), 
    divide_1000
  )
```

```{r results=FALSE, warning=FALSE, message=FALSE}
historical_ladders
```


```{r results=FALSE, warning=FALSE, message=FALSE}
historical_ladders_change <- 
  historical_ladders %>% 
  group_by(country) %>% 
  filter(year %in% c(min(year), max(year))) %>% 
  mutate(dif_ladder = ladder_score - lag(ladder_score, default = first(ladder_score))) %>% 
  mutate(first_year = lag(year, default = first(year))) %>% 
  filter(year == max(year)) %>% 
  ungroup()

historical_ladders_change %>% 
  select(country, year, first_year, dif_ladder) %>% 
  arrange(desc(dif_ladder)) %>% 
  top_n(10)

historical_ladders_change %>% 
  select(country, year, first_year, dif_ladder) %>% 
  arrange(dif_ladder) %>% 
  head(10)
  
```


```{r, fig.show='hide'}
current_ladders_region <- 
  current_ladders %>% 
  select(country, region)

historical_ladders %>% 
  left_join(current_ladders_region, by = "country") %>% 
  group_by(region, year) %>% 
  summarize(mean_ladder = mean(ladder_score)) %>% 
  filter(!is.na(region)) %>% 
  ggplot(aes(year, mean_ladder, color = region)) +
  geom_point() +
  geom_line() +
  theme(legend.position="bottom", legend.box = "horizontal") +
  guides(col = guide_legend(nrow = 2))
```


