---
title: "Exploring Bitcoin"
author: "Avery Rogers"
date: '2021-09-11'
output: pdf_document
categories: []
tags: []
slug: exploring-bitcoin
---

```{r setup, include=FALSE}
# Settings for chunks
knitr::opts_chunk$set(echo = FALSE, results = FALSE, warning = FALSE)

# Libraries
library(tidyverse)
library(lubridate)
library(sqldf)
library(httr)
library(jsonlite)
library(stringr)
library(rvest)
library(Quandl)

```


```{r results=FALSE, warning=FALSE, message=FALSE}
Quandl.api_key('13u77dMYTusnMgECS1zU')
market_price_usd <- Quandl('BCHAIN/MKPRU')
usd_exchange_trade_volume <- Quandl('BCHAIN/TRVOU')
my_wallet_num_users <-Quandl('BCHAIN/MWNUS')
avg_block_size <- Quandl('BCHAIN/AVBLS')
median_transaction_time <- Quandl('BCHAIN/ATRCT')
miners_revenue <- Quandl('BCHAIN/MIREV')
cost_per_transaction <- Quandl('BCHAIN/CPTRA')
transaction_vol_usd_value <- Quandl('BCHAIN/ETRVU')
transaction_vol_value <- Quandl('BCHAIN/ETRAV')
num_transactions <- Quandl('BCHAIN/NTRAN')
mkt_cap <- Quandl('BCHAIN/MKTCP')
total_bitcoins <- Quandl('BCHAIN/TOTBC')
my_wallet_per_day <- Quandl('BCHAIN/MWNTD')


bitcoin_info <- 
  market_price_usd %>% 
  rename(mkt_price_usd = Value) %>% 
  left_join(usd_exchange_trade_volume, by = "Date") %>% 
  rename(usd_exchange_trade_volume = Value) %>% 
  left_join(my_wallet_num_users, by = "Date") %>% 
  rename(my_wallet_num_users = Value) %>% 
  left_join(avg_block_size, by = "Date") %>% 
  rename(avg_block_size = Value) %>% 
  left_join(median_transaction_time, by = "Date") %>% 
  rename(median_transaction_time = Value) %>% 
  left_join(miners_revenue, by = "Date") %>% 
  rename(miners_revenue = Value) %>% 
  left_join(cost_per_transaction, by = "Date") %>% 
  rename(cost_per_transaction = Value) %>% 
  left_join(transaction_vol_usd_value, by = "Date") %>% 
  rename(transaction_vol_usd_value = Value) %>% 
  left_join(transaction_vol_value, by = "Date") %>% 
  rename(transaction_vol_value = Value) %>% 
  left_join(num_transactions, by = "Date") %>% 
  rename(num_transactions = Value) %>% 
  left_join(mkt_cap, by = "Date") %>% 
  rename(mkt_cap = Value) %>% 
  left_join(total_bitcoins, by = "Date") %>% 
  rename(total_bitcoins = Value) %>% 
  left_join(my_wallet_per_day, by = "Date")  %>% 
  rename(my_wallet_per_day = Value) 
  
```

```{r results=FALSE, warning=FALSE, message=FALSE}
bitcoin_info <- 
  bitcoin_info %>% 
  filter(Date < "2021-09-12") %>% 
  rename(date = Date)
```

*Code can be found on Github [right here.](https://github.com/averyr2/r_averyblog/blob/master/content/post/2021-09-11-exploring-bitcoin/index.Rmd)

I will admit that I'm historically something of a crypto skeptic - blame a mix of risk aversion and Silicon Valley's grandiose delusions. But as passes its 12th anniversary, I have come to accept that crypto is here to stay. As a result, I figured I might as well understand its meteoric rise and figure out where it's headed. 

First, I look at Bitcoin's rise, along with the rise in miner revenue. Hindsight is painful, per usual. 

```{r results=FALSE, warning=FALSE, message=FALSE}
bitcoin_info %>% 
  filter(year(date) > 2015) %>% 
  arrange(date) %>% 
  mutate(mkt_price_growth = mkt_price_usd / first(mkt_price_usd) - 1) %>% 
  mutate(miners_revenue_growth = miners_revenue / first(miners_revenue) - 1) %>% 
  select(date, mkt_price_usd, mkt_price_growth, miners_revenue, miners_revenue_growth) %>% 
  ggplot(aes(date)) +
  geom_line(aes(y = mkt_price_growth, color = "Market Price Growth")) +
  geom_line(aes(y = miners_revenue_growth, color = "Miners' Revenue Growth")) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  
  labs(
    x = "Date",
    y = "Percent change in Value from Jan 1, 2016",
    title = "The epic rise of Bitcoin",
    subtitle = "Bitcoin's value has risen over 10,000% since 2016, miner revenue up 2,500%",
  ) +
  theme(legend.position="bottom") +
  theme(legend.title=element_blank())

```

We see that it was better to be an investor than a miner from 2016 to 2021 in terms of percentage growth, although I regret not being either! 

**More data analysis to come soon** 
